---
title: "p8105_hw1_jf3285"
author: Jiarui Fu
date: 2019-09-20
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

## Problem 1
The purpose of this problem is to address variable types and introduce coercion.
<br />Create a data frame comprised of:
a random sample of size 8 from a standard Normal distribution, 
a logical vector indicating whether elements of the sample are greater than 0, 
a character vector of length , 
a factor vector of length 8, with 3 different factor “levels”.
```{r df_create}
# set seed to ensure reproducibility
set.seed(1234)

# create data frame
create_df = tibble(
  x = rnorm(8),
  vec_logical = x > 0,
  vec_char = c("My", "name", "is", "Alexa", "Hello", "Hey", "Hi","Bye"),
  vec_factor = factor(c("low", "medium", "high", "high", "medium", "low", "medium", "medium"))
)

# calculate the mean of each variable
# numeric and logical vectors work but character and factor vectors return NA
x_mean = mean(pull(create_df, x))
vec_logical_mean = mean(pull(create_df, vec_logical))
# vec_char_mean = mean(vec_char)
# vec_factor_mean = mean(vec_factor)

as.numeric(pull(create_df, vec_logical))
# as numeric(character vector) returns NA
as.numeric(pull(create_df, vec_char))
as.numeric(pull(create_df, vec_factor))
```
This part of code chunk converts logical vectors to numeric and factor vectors.
```{r convert}
# logical vectors convert to numeric and factor vectors
convert_logi_numeric = as.numeric(pull(create_df, vec_logical))
result1 = convert_logi_numeric*(pull(create_df, x))
result1
convert_logi_factor = as.factor(pull(create_df, vec_logical))
result2 = convert_logi_factor*(pull(create_df, x))
result2
double_convert = as.numeric(convert_logi_factor)
result3 = double_convert*(pull(create_df, x))
result3
```

## Problem 2
This problem focuses on the use of inline R code and plotting. 
```{r df_plot}
set.seed(1234)

# create data frame
plot_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  logi_vec = c(x+y) > 1,
  num_vec = as.numeric(logi_vec),
  factor_vec = as.factor(logi_vec)
)

# make and export scatterplot
ggplot(plot_df, aes(x=x, y=y, color=logi_vec)) + geom_point()
ggsave("scatter plot.pdf", height = 4, width = 6)
ggplot(plot_df, aes(x=x, y=y, color=num_vec)) + geom_point()
ggplot(plot_df, aes(x=x, y=y, color=factor_vec)) + geom_point()
```

# Description of vectors using inline R code
 The size of the dataset is `r nrow((plot_df))` and `r ncol((plot_df))`.
 
 The mean of x is `r mean(pull(plot_df, x))`.
 
 The median of x is `r median(pull(plot_df, x))`.
 
 The standard deviation of x is `r sd(pull(plot_df, x))`.
 
 The proportion of cases for which x + y > 1 is `r sum(pull(plot_df, logi_vec), na.rm = TRUE)/500`
 






